### 2021-04-05 Chapter-4/Chapter-4.tex
Updates from Overleaf

Since the table look-ups and conditional jumps are replaced with single-bit logical gates, with no secret-dependent memory addresses or control flow, both the data access and control flow types of side-channel leakages are mitigated.
\section{Discussion}
While recent work found many side-channel vulnerabilities,
we note that many of them have not been patched by developers.
Side-channels are ubiquitous in software and it would be difficult to fix all of them.
We need a tool that estimates the sensitivity of each vulnerability
so software engineers can focus on
``severe'' leakages. For example, \tool{} reports that
the modular exponentiation using square and multiply algorithms can
leak more information than a key validation function.

>>>>

Since the table look-ups and conditional jumps are replaced with single-bit logical gates, with no secret-dependent memory addresses or control flow, both the data access and control flow types of side-channel leakages are mitigated.
\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
uint8_t password = input();

uint8_t SBOX[] = {1, 0, 3, 1, 2, 2, 3, 0};

if (password <= 0b111)      \\Leaks 5 bits of password
    ret = SBOX[password];   \\Leaks 4 bits of password
      \end{lstlisting}
    \caption{SBOX without BitSlicing}
    \label{fig:SBOX_da}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
uint8_t password = input();

a = *password & 0b001;
b = (*password & 0b010) >> 1;
c = (*password & 0b100) >> 2;

na = ~a & 1;
nb = ~b & 1;
nc = ~c & 1;

t0 = (b & nc);
t1 = (b | nc);

l = (a & nb) | t0;
r = (na & t1) | t0;

ret = l << 1 + r;
\end{lstlisting}
\caption{SBOX with BitSlicing}
\label{fig:SBOX_bitslicing}
\end{figure}
We test \tool\ on bit-slicing. We adopt the SBOX
implementations, commonly used in block ciphers such as DES and AES, with and
without bit-slicing, and apply \tool{} to confirm the mitigation. The SBOX implementation with and without bit-slicing are shown in Figure~\ref{fig:SBOX_bitslicing} and Figure~\ref{fig:SBOX_da}, respectively. Considering an SBOX take some bits derived from a password as input and output 2-bit transform result, the plain implementation
has a range check on the password input and a secret-dependent table lookup while bit-slicing does not.

\tool\ reports that there are control-flow and data access types of leakages
in the non-bit-slicing implementation (line 6 and line 7 in Figure~\ref{fig:SBOX_da}). The number of leaked bits is
5.0 and 4.4, respectively. At line 6, according to Definition~\ref{def}, the
input set $K$ is $[0,2^8-1]$, the observed input set $K^o$ is $[0,2^3-1]$. Thus,
the leakage $L_{\beta(k)\rightarrow o}$ based on the observation ($o$) is
$L_{\beta(k)\rightarrow o} = \log_2{|K|} - \log_2{|K^o|} = 8-3 = 5$ bit, which
confirms the result from \tool. Similarly, we can verify the result of the other
leakage site. \tool{} reports no leakage on the bit-slicing implementation.

\subsubsection{Smaller Lookup Tables}
\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
static const uint8_t T[16] = {
      0x63U, 0x7cU, 0x77U, 0x7bU, 0xf2U, 0x6bU, 0x6fU, 0xc5U,
      0x30U, 0x01U, 0x67U, 0x2bU, 0xfeU, 0xd7U, 0xabU, 0x76U};
      \end{lstlisting}
    \caption{One-byte Entry Lookup Table}
    \label{fig:one_byte_table}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
static const uint32_t T[16] = {
    0xc66363a5U, 0xf87c7c84U, 0xee777799U, 0xf67b7b8dU,
    0xfff2f20dU, 0xd66b6bbdU, 0xde6f6fb1U, 0x91c5c554U,
    0x60303050U, 0x02010103U, 0xce6767a9U, 0x562b2b7dU,
    0xe7fefe19U, 0xb5d7d762U, 0x4dababe6U, 0xec76769aU};
    \end{lstlisting}

    \caption{Four-byte Entry Lookup Table}
    \label{fig:four_byte_table}
\end{figure}
\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
void encrypt_one(uint32_t *o, uint32_t *key, uint32_t l) {
    for (int i = 0; i < l; i+=4)
        output[i] = (T[(key[i]>>24)] << 24) ^
                    (T[(key[i+1]>>16) & 0xff] << 16) ^
                    (T[(key[i+2]>>8) & 0xff] << 8) ^
                    (T[(key[i+3]) & 0xff]);
}
    \end{lstlisting}
    \caption{Encrypt with One-byte Entry Lookup Table}
    \label{fig:one_byte_table_lookup}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.02\textwidth,xrightmargin=.01\textwidth]
void encrypt_four(uint32_t *o, uint32_t *key, uint32_t l) {
    for (int i = 0; i < l; i+=4)
        output[i] = (T[(key[i]>>24)] & 0xff000000) ^
                    (T[(key[i+1]>>16) & 0xff] & 0x00ff0000) ^
                    (T[(key[i+2]>>8) & 0xff] & 0x0000ff00) ^
                    (T[(key[i+3]) & 0xff] & 0x000000ff);
}
    \end{lstlisting}
    \caption{Encrypt with One-byte Entry Lookup Table}
    \label{fig:four_byte_table_lookup}
\end{figure}

Considering the cache-collision timing attack~\cite{Bonneau11894063_16}, the
probability of leakage decreases when the lookup table entry or element size gets
smaller. We take AES as an example, for the last round of encryption, the algorithm
uses a specialized table with 4-byte size entries, of which only one
byte from each element actually contributes to the result. Hence, a significant method to reduce
the information leakage in AES is to reduce the size of the lookup table. We
designed a demo to illustrate this countermeasure, a lookup table with one-byte
size elements (see Figure~\ref{fig:one_byte_table}) will leak less information
than a table with 4-byte element size (see Figure~\ref{fig:four_byte_table}). 
The encrypt function for each table to process the same amount of information, 
respectively, (see Figure~\ref{fig:one_byte_table_lookup} and
Figure~\ref{fig:four_byte_table_lookup}).  \tool{} reports three leakage sites for
each lookup table, with 4.0 bits for the table with smaller entries and 2.0
bits for the table with bigger entries, confirming the theory.

\section{Discussion}
While recent work found many side-channel vulnerabilities,
we note that many of them have not been patched by developers.
Side-channels are ubiquitous in software and it would be difficult to fix all of them.
We need a tool that estimates the sensitivity of each vulnerability
so software engineers can focus on
``severe'' leakages. For example, \tool{} reports that
the modular exponentiation using square and multiply algorithms can
leak more information than a key validation function.
